\documentclass[a4paper,12pt]{article}
\usepackage[utf8x]{inputenc}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{subfig}
\usepackage[countmax]{subfloat}
\newcommand{\tab}{\hspace*{2em}}
%opening

\begin{document}


\begin{center}
{\Large Previous Research}

<<<<<<< HEAD
\emph{Matthew Le}
\end{center}

I am currently working on two research projects, both of which are funded by the NSF Research Experience for Undergraduates (REU) program. The first project is under the guidance of Professor Eric Van Wyk, dealing with finding ways to design programming languages, such that they can easily be extended, to support additional constructs. The second project that I have been working on deals with investigating the effects of Pacific Ocean warming on Atlantic hurricane activity. For this project, I have been working under the mentorship of Professor Vipin Kumar while working closely with one of his graduate students and a senior research scientist in Professor Kumar's laboratory.

Domain specific languages (e.g. R, MATLAB, SQL, \textit{etc.}) have become popular because they allow the programmer to write code at a high level of abstraction using constructs relevant to their domain, while enjoying the benefits of domain specific optimizations. The downside of creating a new language to add these domain specific features is that you lose portability. Also, since it is a different language, previously existing code must be rewritten in the new language. Our research aims to find ways to implement these domain specific features as language extensions to already existing general purpose languages. This way, previously written code can still be utilized, and the extensions can be used on all architectures that the host language is built for.  
=======
I am currently working on two research projects, both of which are funded by the NSF Research Experience for Undergraduates (REU) program. The first project is under the guidance of Professor Eric Van Wyk, dealing with finding ways to design programming languages, such that they can easily be extended [extended to do what? for what?]. The second project that I have been working on deals with investigating the effects of Pacific Ocean warming on Atlantic hurricane activity. For this project, I have been working under the mentorship of Prof. Vipin Kumar while working closely with a graduate student and a senior research scientist in Prof. Kumar's laboratory.  

Domain specific languages (e.g mobile computing, genomics, \emph{etc.}) have become popular because they allow the programmer to write code at a high level of abstraction, while enjoying the benefits of domain specific optimizations. The downside of creating a new language to add these domain specific features is that you lose portability as such specialized languages cannot be generalized to other domains. Also, since it is a different language with different syntax, existing code must be rewritten in the new language [in what context do you need to use existing code that must be re-written?]. Our research aims to find ways to implement these domain specific features as language extensions to already existing general purpose languages. This way, previously written code can still be utilized, and the extensions can be used on all platforms [what do you mean by platforms in this context?] that the host language is built for.  
>>>>>>> my version of edits

The project that I am working on implements this idea by first creating a host language. For this, we have chosen a prototype language, which is a subset of C. The role of the host language is to show the state of a general purpose programming language in its initial state without any extensions.  We chose to use a subset of C because its low level of abstraction gives rise to programming complexity, but when used correctly it can yield code that runs very efficiently. After the host language was implemented, we began working on extensions. The extension that I have been working on adds support for matrices, and the operations that accompany them. We create these extensions using a technique known as ``forwarding'', which works by reducing extensions to constructs that already exist in the host language.  The extension that I am working on makes heavy use of forwarding, such that this extensions is ``pure'', meaning it is completely reducible to constructs existing in the host language.  The benefit to this is that if the host language gets updated, then the extension does not need to be modified.  Also, this allows for other language extensions to be simultaneously and independently implemented.

I began implementing the extension by first designing the underlying representation of matrices. Since we want this matrix extension to generalize to any number of dimensions, we chose to represent matrices as a structure containing the size of each dimension, and the data the matrix contains. We then index matrices using row-major indexing. After designing the underlying representation, we created an additional type for matrices, which then forwards to a host type, in this case we forward to the host struct type. We then created syntax and abstract productions for declaring matrices.  Matrix declarations are really no different than declaring any other variable, we simply need to add the name of the matrix being declared to the symbol table. Another necessity that we implemented was matrix initialization. This production forwards to a sequence of assignment statements, where we allocate space for the structure, the dimension sizes array, and the data array. A few other productions that we added to the core matrix extension include matrix references, matrix arithmetic, and printing of matrices.  

After creating an extension that simply introduces this new data structure and some very basic operations, we started adding some of the domain specific features that are found in MATLAB.  Such additional functions include logical indexing, referencing multiple elements of a matrix at a time, assigning to multiple elements of a matrix, and matrix permutations.  We ended up putting this functionality into an extension of its own, showing that not only host languages can be extended in a modular way, but extensions can also be further extended in a modular way.  These additional features have presented many opportunities for generating parallel C code, which is something that we are currently exploring now.

As of now, we currently have a functioning language extensions that supports just about everything that one would be able to do in MATLAB. We are able to generate C code from this extension, and are currently working on the optimization phase of the project. I have implemented many of the algorithms that we use in the data mining research lab that I am also apart of, using the language extension that I have created. The two versions of most algorithms run in about the same amount of time, and produce the same results. The fact that these two versions run in a similar amounts of time, without any optimizations is very encouraging, because this implies that once we are able to generate parallel code, we will have a large runtime advantage over MATLAB on top of the fact that we have successfully integrated this extension into a general purpose language.

A second research opportunity that I am currently working on is another NSF REU funded project in the field of climate data mining, under the guidance of graduate student James Faghmous and his advisor Professor Vipin Kumar. The El-Ni\~{n}o Southern Oscillation (ENSO) is a phenomenon where Pacific Ocean warming has been linked to Atlantic hurricane activity\cite{enso}.  The problem is that the correlation between the two is quite poor.  This research effort aims to further investigate this relationship. 

When I began working on this project, James had already developed a method that explained the Pacific-Atlantic relationship more accurately than ENSO. Traditionally, climate scientists 
have expressed the impact of the Pacific ocean on Atlantic hurricanes by monitoring static oceanic regions and recording their sea surface temperature. James proposed a method that instead of monitoring the temperature of the Pacific Ocean, one can monitor the location of the warming, i.e. what subset of the Pacific is the warmest. After constructing an index based on the location of the warming instead of its absolute value, we were able to develop an index that is more intimately related to Atlantic Hurricane activity. After finalizing the original index, I went on to experiment with variations of the index to increase its accuracy and scientific interpretability. Many of these variations involved incorporating additional variables in addition to sea surface temperature.  After a few months, I had come up with an index of my own that performed better than the previous one with an increase from 0.6 to 0.8 correlation between the index and annual Atlantic hurricane counts. This index has remained our top performing predictor of Atlantic hurricane activity when we restrict ourselves to the Pacific Ocean. Over the past few months we have been compiling our results and subjecting the index to rigorous statistical significance testing. Given the relative short observational record (~30 years of data) and the auto-correalted nature of climate data (data that are close in space and time tend by related), we could not use traditional statistical significance tests (e.g. t-test) to ensure that the high correlation between our index and Atlantic hurricane is not due to random chance. Instead, I helped implement a range of randomization studies that attempt to re-create a null distribution based on the observed data and we used that distribution to quantify the significance of our results.




<<<<<<< HEAD
In addition to conducting research, I have also had a few experiences with presenting my results.  About four months into working on my programming languages research project, we presented at the University of Minnesota Undergraduate Research Symposium.  this was an especially interesting experience because we were not just presenting to computer scientists.  It proved to be a rather difficult task to explain what we were working on to people who had never been exposed to computer science before. I also had the opportunity to present some of our work in the data mining research lab at the Expeditions in Computing Workshop hosted by the University of Minnesota. 
=======



%The research group deals with using data mining and machine learning techniques to further understand the causes and effects of climate change.  The particular project that I have been working on deals with discovering the effects of Pacific Ocean warming on Atlantic Ocean hurricane activity.  When I first began working on the project, they already had developed an algorithm that tracks the location of the region that has the warmest sea surface temperature, and used these metrics to predict hurricane activity in the Atlantic.  They found that if the location of the warmest region of the Pacific is in the western portion of the ocean, then there tends to be more hurricane activity than if the warmest region is in the eastern portion of the ocean.  The correlations of this new index proved to be much more accurate in predicting hurricane activity than traditional ENSO indices, which simply look at the sea surface temperature of a static region in the Pacific\cite{enso}.\newline
%\tab After a few months of working on the project, I developed a new index, which incorporates other variables into the mix, but still maintains this idea of monitoring spatial distributions.  I began working on this new index when I noticed that other indices that also tried to predict hurricanes used other variables such as sea level pressure and outgoing long wave radiation.  I tried using the same method that we used in the previous index on some of these new variables, as well as simply taking the magnitude of the variables within various regions.  I got some decent results, but nothing substantial.  After a while I attempted to combine these various indices into one index using various linear combinations, however, there was always this problem that I was trying to combine variables that had different units, rendering my results as nonsense.  I eventually realized that if I were to take the z-score of these indices, I could put everything into relative terms.  This approach ended up working very well, and the results that we saw from this were much better than the results that we were getting previously. We are currently in the process of compiling our results and preparing this work for publication. \newline

In addition to conducting research, I have also had a few gratifying experiences presenting my results. About four months into working on my programming languages research project, we presented at the University of Minnesota Undergraduate Research Symposium. This was an interesting experience because we were not just presenting to computer scientists. It proved to be a rather difficult task to explain what we were working on to the general public who had never been exposed to computer science. I am know more thoughtful about my communication as I value the important to communicating the results of federally funded research to the broader community. I also had the opportunity to present a poster on my work at the annual NSF Expeditions Grant workshop at the University of Minnesota, where I interacted with a broad range of scientists, faculty, and graduate students from over 10 universities. 
>>>>>>> my version of edits

I have had a great time working on these two REU projects, and feel that they have stimulated my interest in conducting research in the future.  I certainly have a different perspective on research now that I have had some exposure to it, and feel that I am much more prepared than previously in carrying out research at the graduate level.  
\begin{thebibliography}{9} \vspace{-1ex}
\footnotesize
\bibitem{enso}M.C. Bove, J.J. O'Brien, J.B. Elsner, C.W. Landsea, X. Niu.  Effect of El Nino on U.S. Landfalling Hurricanes, Revisted.  \textit{Bulletin of the American Meteorological Society}, Volume 79, Number 11
\end{thebibliography}
\end{document}










